
# Additional network properties (PPI subsets, assorativity)

## Properties and performance of PPI for large-scale and curated database


```{r}
# read data
library(pacman)

if(!file.exists("../cache/PPI_subset_and_CoexCore_properties.RDS")){
  source("../source/compute_properties_PPIsubsets_and_CoexCore.R")
}

g_prop_Labelled_df <- readRDS("../cache/PPI_subset_and_CoexCore_properties.RDS")
```

```{r plot the results, fig.width=9, fig.height=2.5,  layout="l-body-outset"}
p_load(ggforce, patchwork)

# create a list of  plots to patch together
plot_combine = list()
 for(type_network in unique(g_prop_Labelled_df$type)){
  plot = list()
    for(prop in unique(g_prop_Labelled_df$property)){
    plot[[prop]] = g_prop_Labelled_df  %>% 
      filter(type == type_network) %>%
      arrange(group) %>% filter(property == prop) %>%
      ggplot( aes(x=group, y=as.numeric(value))) + 
    #  facet_grid(type ~ ., scales = "free", space = "free") +
      geom_segment( aes(x=group, xend=group, y=0, yend=value), color="grey80", size=1.5) +
      geom_violin(fill="#F8B100", alpha = 0.4, color = NA) +
      geom_point( aes(color=alphaval), size=4, alpha=0.6) +
      geom_sina(aes(color='0'), size=4, alpha=0.6) + # add jitters
      geom_hline(yintercept=0, color='grey50', size=1,  linetype='dotted') +
      theme_light() + 
      coord_flip() +
      theme(
        panel.grid.major.y = element_blank(),
        panel.border = element_blank(),
        axis.ticks.y = element_blank(),
       # axis.text.y = element_blank(),
      ) +
      guides(color = F)+
      xlab("") +
      scale_color_manual(values = c("#F8B100",NA)) +
      ylab(prop)
    
    # scale y log for some properties (n edges)
    if(prop %in% c("Number of edges")){
      plot[[prop]] = plot[[prop]] + scale_y_log10()
    }
    
    # for the first plot, allows axis label
    if(!prop %in% c("Number of nodes")){
      plot[[prop]] = plot[[prop]] +  theme(axis.text.y = element_blank())
    }
    
  }
  
  plot_combine[[type_network]] = plot$`Number of nodes` + plot$`Edge density` + plot$`Global clustering` + plot$Assortativity + plot$`Social bias` + plot_layout(nrow = 1)

}


plot_combi <- plot_combine$PPI/plot_combine$`Co-expression`


# uncomment to save the plot as pdf
#ggsave("../Figs/network_properties_characterisation.pdf", plot_combine, height = 2.5, width = 9)

suppressWarnings(print(plot_combi))

#ggsave("../Figs/additional_network_properties.pdf", plot = plot_combi, height = 4, width = 10)
```

## Assortativity (Suppl. Fig. 3e-f)

- Average deegree of neighbours ($k_{nn}$)
- Local assortativity

Two methods additionally employed to test dissortativity and prove its derivatives from noth high-throughput and curated PPI

```{r knn}
sumstat <- citation_assortativity %>% group_by(inBioPlexbait) %>% summarise(summary = list(summary(neighbour_degree)))
sumstat[1,"summary"][[1]]
sumstat[2,"summary"][[1]]

p_count_neighbour <- ggplot(citation_assortativity, aes(x = count, y = neighbour_degree)) + 
  geom_point(alpha = 0.1, col = "grey75") + theme_cowplot() + 
  geom_point(data = citation_assortativity %>% 
               group_by(count) %>% 
               summarise(mean = mean(neighbour_degree)), 
             aes(x = count, y = mean), col = "red", shape = 4)  + 
  scale_y_log10() + 
  scale_x_log10() +
  xlab("# PubMed Count") + 
  ylab(expression("<"~italic(k[nn])~"("~italic(k)~")>")) + 
  ggtitle("PubMed count vs its PPI neighbours", 
          subtitle = paste0("Pearson's r:", round(cor(citation_assortativity %>% 
                                                        filter(!is.na(count), !is.na(neighbour_degree)) %>% pull(count), citation_assortativity %>% 
                                                        filter(!is.na(count), !is.na(neighbour_degree)) %>% pull(neighbour_degree)), 3)))



p_degree_neighbour <- ggplot(citation_assortativity, aes(x = degree, y = neighbour_degree)) + 
  geom_point(alpha = 0.1, col = "grey75") + 
  theme_cowplot() + 
  geom_point(data = citation_assortativity %>% 
               group_by(degree) %>% 
               summarise(mean = mean(neighbour_degree)), 
             aes(x = degree, y = mean), col = "red", shape = 4)  + 
  scale_y_log10() + 
  scale_x_log10() +
  xlab(expression(italic(k))) + ylab(expression("<"~italic(k[nn])~"("~italic(k)~")>")) + 
  ggtitle("PPI degrees vs its neighbours", 
          subtitle = paste0("Pearson's r:", round(cor(ppi_degree, neighour_degree_avg), 3)))

p_assortativity <- p_count_neighbour + p_degree_neighbour
#ggsave("../Figs/local_assortativity_PPI.pdf", plot = p_assortativity, width = 8, height = 3)

p_assortativity
```

```{r}
bioplex_baits_293T <- read_tsv("../data/raw_data/Bioplex_293T-baits.tsv")
bioplex_baits_293T$cell = "293T"
bioplex_baits_HCT116 <- read_tsv("../data/raw_data/Bioplex_HCT116-baits.tsv")
bioplex_baits_HCT116$cell <- "HCT116"

bioplex_baits <- rbind(bioplex_baits_293T, bioplex_baits_HCT116)


ppi_degree <- degree(g$PPI_HIPPIE)
ppi_degree = ppi_degree[ppi_degree>0]

p_k = degree_distribution(g$PPI_HIPPIE)
p_k = p_k[-1]

# q_k =  remain degree distribution
q_k = (1:max(ppi_degree) +1)*c(p_k[-1],0)/mean(ppi_degree)
#q_k = ppi_degree-1

mean_q  = sum(1:max(ppi_degree)*q_k)
variance_q = sum((mean(ppi_degree)-1:max(ppi_degree))^2 * q_k)


neighour_degree_avg = knn(g$PPI_HIPPIE)$knn 
neighour_degree_avg = neighour_degree_avg[names(ppi_degree)]

local_assortativity = (ppi_degree-1)*ppi_degree*(neighour_degree_avg - 1 - mean_q)/(2*ecount(g$PPI_HIPPIE)*variance_q)

cuts <- c(0,1,2,5, 10,20, 50,100, 200, 500,1000,5000)
citation_assortativity = full_join(citation_count,
  tibble(gene = names(ppi_degree), 
         local_assortativity = local_assortativity,
         degree = ppi_degree,
         neighbour_degree = neighour_degree_avg,
         degree_binned = cut(degree, breaks = cuts, labels = cuts[-1]),
         label = ifelse(local_assortativity < -0.002, gene, ""))) %>%
  filter(!is.na(local_assortativity)) %>%
  mutate(inBioPlexbait = gene %in% bioplex_baits$`Bait Symbol`)


library(cowplot)

p_local_assortativity <- ggplot(citation_assortativity, aes(x = degree, y = local_assortativity, col = inBioPlexbait))+ 
  geom_point(alpha = 0.4) + 
  scale_x_log10() + 
  theme_cowplot()  + 
  guides(col = F) + 
  xlab("Degree") + 
  ylab("Local assortartivity") + 
  ggrepel::geom_text_repel(aes(label = label)) + 
  scale_color_brewer(type = "qual", direction = -1) + 
  facet_grid(.~inBioPlexbait)

#ggsave("../Figs/local_assortativity_colored_by_bioplex.pdf", plot = p_local_assortativity, width = 4, height = 3)
p_local_assortativity
```

```{r}
#ggplot(citation_assortativity %>% mutate(local_assortativity_cut = cut(local_assortativity, breaks = c(-1,-0.002,-0.001,-0.0005,-0.0001, -0.00005, -0.000001, 0, 1))) , aes(x = local_assortativity_cut, fill = inBioPlexbait))+ geom_bar(position = "dodge") + theme_cowplot()  + guides(col = F) #+ xlab("Degree") + ylab("Local assortartivity") + ggrepel::geom_text_repel(aes(label = label))

#ggplot(citation_assortativity, aes(x = degree_binned, fill = inBioPlexbait))+ geom_bar(position = "dodge") + theme_cowplot()  + guides(col = F) #+ 


p_local_assortativity_bioplex <- ggplot(citation_assortativity,  aes(x = inBioPlexbait, y= neighbour_degree, fill = inBioPlexbait)) + 
  geom_boxplot()  + 
  theme_cowplot()  + 
  guides(col = F) + 
  scale_y_log10() + 
  ggpubr::stat_compare_means() + 
  theme_cowplot()  + 
  guides(fill = F) + 
  scale_fill_brewer(type = "qual", direction = -1) + 
  xlab("BioPlex baits") + ylab("Average neighbour degree")

#ggsave("../Figs/local_assortativity_colored_by_bioplex_box.pdf", plot = p_local_assortativity_bioplex, width = 4, height = 3)
p_local_assortativity_bioplex
```

## Bioplex
```{r}
g_bioplex_293T <- read_tsv("../data/raw_data/Bioplex_293T-interactions.tsv", col_types = "-c-c") %>% graph_from_data_frame(.,directed = "F")
g_bioplex_HCT116 <- read_tsv("../data/raw_data/Bioplex_HCT116-interactions.tsv", col_types = "-c-c") %>% graph_from_data_frame(.,directed = "F")


assortativity_nominal(g_bioplex_293T, types =  ifelse(V(g_bioplex_293T)$name %in% bioplex_baits_293T$`Bait Symbol`,1, 2), directed = F)
assortativity_nominal(g_bioplex_HCT116, types =  ifelse(V(g_bioplex_HCT116)$name %in% bioplex_baits_HCT116$`Bait Symbol`,1, 2), directed = F)


annotate_assortativity = function(graph){
  
    g_degree <- degree(graph)
    g_degree = g_degree[g_degree>0]
    
    p_k = degree_distribution(graph)
    p_k = p_k[-1]
    
    
    q_k = (1:max(g_degree) +1)*c(p_k[-1],0)/mean(g_degree)
    #q_k = ppi_degree-1
    
    mean_q  = sum(1:max(g_degree)*q_k)
    variance_q = sum((mean(g_degree)-1:max(g_degree))^2 * q_k)
    
    neighour_degree_avg = knn(graph)$knn 
    neighour_degree_avg = neighour_degree_avg[names(g_degree)]
    
    local_assortativity = (g_degree-1)*g_degree*(neighour_degree_avg - mean_q)/(2*ecount(graph)*variance_q)
    

    citation_assortativity = full_join(citation_count,
      tibble(gene = names(g_degree), 
             local_assortativity = local_assortativity,
             degree = g_degree,
             neighbour_degree = neighour_degree_avg,
             label = ifelse(local_assortativity < -0.002, gene, ""))) %>%
      filter(!is.na(local_assortativity)) 

  return(citation_assortativity)
}


citation_assortativity_293T <- annotate_assortativity(g_bioplex_293T)
citation_assortativity_293T <- citation_assortativity_293T %>%
  mutate(cell = "293T",
         inBioPlexbait = gene %in% bioplex_baits_293T$`Bait Symbol`)
citation_assortativity_HCT116 <- annotate_assortativity(g_bioplex_HCT116) %>%
  mutate(cell = "HCT116",
         inBioPlexbait = gene %in% bioplex_baits_HCT116$`Bait Symbol`)

citation_assortativity_bioplex = rbind(citation_assortativity_HCT116, citation_assortativity_293T) %>%
  mutate(local_assortativity_cut = cut(local_assortativity, breaks = c(-1,-0.004,-0.002,0,1)),
         inBioPlexbait = factor(inBioPlexbait, levels = c(F,T), labels = c("non bait", "baits")))

p_bioplex_cell_assortaticity <- ggplot(citation_assortativity_bioplex, aes(x = degree, y = local_assortativity, col = inBioPlexbait))+ geom_point() + scale_x_log10() + theme_cowplot()  + guides(col = F) + xlab("Degree") + ylab("Local assortartivity")  + facet_grid(cell~inBioPlexbait) + scale_color_brewer(type = "qual", direction = -1) 

#ggsave("../Figs/local_assortativity_colored_by_bioplex_bioplexNet.pdf", plot = p_bioplex_cell_assortaticity, width = 4, height = 3)

p_bioplex_cell_assortaticity
```


## Disease associaton of PPI subsets

```{r}

p_load(ggforestplot)
rare_genetic_result_folder = "../cache/output/Orphageneset_rare_additionalNetworks//"
rare_genetic_heatmap_file = "../cache/heatmap_network_disease_association_orphanets_genetic_rare_diseases_additionalNetworks.pdf"


# for printing (cm)
A4width = 21
A4height = 29.7


# defines colours used in the heatmap
cols = RColorBrewer::brewer.pal(5, "Blues")
cols = cols[2:5]

# load  network labels
network_info = read_tsv("../data/network_details.tsv")
source("../functions/process_LCC_result.R")


  ## process the results LCC significance
  result_df = readRDS(paste0(rare_genetic_result_folder, "LCC_and_distance_calculation_results.RDS"))
  
  processed_result_df = process_LCC_result(result_df)
  
  processed_result_df <- processed_result_df %>% select(-c(type, main_type, subtype, source)) 
  
  processed_result_df <- left_join(processed_result_df, g_prop_df %>% select(-c(property, value, alphaval)) %>% distinct())
  
  p <- processed_result_df  %>% dplyr::filter(LCC.signif != "none") %>%
    ggplot(aes(name, network)) + geom_tile(fill = "white") + #facet_grid(.~main_type, space = "free", scale = "free")  +## to get the rect filled
    # geom_point(aes(size = N_in_graph*1.7),  colour = LCC.signif)  +   ## geom_point for circle illusion
     geom_stripes(odd = "grey90", even = "#00000000") +
  #  theme_light() + 
   # theme_forest() +
    theme_minimal_hgrid() +
    theme(panel.spacing = unit(0.25, "lines"),
          axis.text.x = element_text(angle = 90, hjust = 1), 
          panel.grid.major.y = element_line(colour="grey", linetype="dotted")) +
      geom_point(aes(
      size = log10(N_in_graph),
      fill = LCC.signif,
      colour = LCC.signif), alpha = 0.7)  + 
    scale_color_manual(values = cols) +
    # scale_size(range = c(1, 10))+             ## to tune the size of circles
      coord_flip() +
    labs(x="", y="") + guides(size = F)
  
  
p
```

It shows that PPI HuRI alone is very sparse to detect disease module in the first place.
