# Characterizing the network architectures across biological scales 


```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(pacman)
p_load(patchwork, igraph, tidyverse, cowplot, rmarkdown)

# compute different network properties
if(!file.exists("../cache/network_complementarity_topological.RDS")){
  # load required functions
source("../source/network_properties_analysis.R")

} else{
  print("Load precomputed data")
  g_prop_df <- readRDS("../cache/network_complementarity_topological.RDS")
}

network_details <- read_tsv("../data/network_details.tsv", col_types = 'ccccc')
```

## Network details (Suppl. Data 1)

45 Network layers from six major databases were constructed as detailed below:

```{r}
paged_table(network_details %>% select(!type))
```


## Topological complementarity (Fig. 1h)

In addition to the scale comprehensiveness, these networks are also topologically complementary. A number of key network properties including node and link coverage, modularity, assortativity, and social bias, have been compared and shown below.

Social bias: many networks were constructed based on curation from literatures. The social bias of a network is assessed by the Spearman’s correlation coefficient between the network degree of a gene and the number of publications mentioning the gene. The number of publications was queried using the INDRA python module (http://www.indra.bio, accessed on 12 April 2019)

```{r}
g_prop_df %>% 
  mutate(type = ifelse(grepl("coex", network), "co-expression",network)) %>%
  group_by(type, property) %>%
  summarise(value = mean(value)) %>%
  paged_table()
```

Note that, for the co-expression layer, the values showed on the table above are averaged from all 38 tissue-specific networks.

The plot below summarises the table properties.

```{r plot the results, fig.width=9, fig.height=2.5,  layout="l-body-outset"}

# create a list of  plots to patch together
plot = list()
for(prop in unique(g_prop_df$property)){
  plot[[prop]] = g_prop_df  %>% 
    arrange(group) %>% filter(property == prop) %>%
    ggplot( aes(x=group, y=as.numeric(value))) + 
    geom_segment( aes(x=group, xend=group, y=0, yend=value), color="grey80", size=1.5) +
    geom_violin(fill="#F8B100", alpha = 0.4, color = NA) +
    geom_point( aes(color=alphaval), size=4, alpha=0.6) +
    theme_light() + 
    coord_flip() +
    theme(
      panel.grid.major.y = element_blank(),
      panel.border = element_blank(),
      axis.ticks.y = element_blank(),
     # axis.text.y = element_blank(),
    ) +
    guides(color = F)+
    xlab("") +
    scale_color_manual(values = c("#F8B100",NA)) +
    ylab(prop)
  
  # scale y log for some properties (n edges)
  if(prop %in% c("Number of edges")){
    plot[[prop]] = plot[[prop]] + scale_y_log10()
  }
  
  # for the first plot, allows axis label
  if(!prop %in% c("Number of nodes")){
    plot[[prop]] = plot[[prop]] +  theme(axis.text.y = element_blank())
  }
  
}

plot_combine = plot$`Number of nodes` + plot$`Edge density` + plot$`Global clustering` + plot$Assortativity + plot$`Social bias` + plot_layout(nrow = 1)

# uncomment to save the plot as pdf
#ggsave("../Figs/network_properties_characterisation.pdf", plot_combine, height = 2.5, width = 9)

suppressWarnings(print(plot_combine))
```

## The Social bias of the networks (Suppl. Fig. 1f)


```{r citation analysis, eval=FALSE, include=FALSE}
# plotting the degree citation, not executed
degree_citation_df <- readRDS("../cache/degree_citation_correlation.RDS")

network_to_show <- c("ppi", "HP", "GOBP", "coex_HRV")

degree_citation_df_selected <- degree_citation_df[network_to_show] %>% 
  bind_rows(, .id = "network") %>%
  mutate(lab = ifelse(rank<5, gene, ""))

pacman::p_load("hexbin")
p <- ggplot(degree_citation_df_selected, aes(y=count, x=degree)) + 
  geom_hex(bins = 40) + 
  scale_fill_gradient(low = "white", high = "navy") +
  geom_smooth(method = 'lm', se = F, color = 'red', linetype = "dashed") +
  #ggrepel::geom_text_repel(aes(label = lab))+
  ggpubr::stat_cor(method = "spearman", p.accuracy = 0.001, r.accuracy = 0.01) +
  facet_grid(.~network) +
  theme_cowplot() + 
  scale_x_log10() + 
  scale_y_log10() +
  ylab("# PubMed") +
  xlab("Network degree")

#ggsave("../Figs/hexbin_plot_network_degree_vs_pubmed_count.pdf", p, width = 9, height = 3)

p
```

##  The network similarity (Fig. 1f, g)


We quantified the similarity of a given pair of networks $g_A \in G(V_A, E_A)$ and $g_B \in G(V_A, E_A)$ using the edge overlap index:
$$S_{AB}=\dfrac{|E_A \cap E_B|}{\text{min}(|E_A|,|E_B|)}$$
We used a dissimilarity measure defined as $d_{AB} = 1 - S_{AB}$ to construct a 2D map $\mathbf{X} \subset \mathbb{R}^{2}$ that preserves network dissimilarities by employing Kruskal’s non-metric multidimensional scaling (R package MASS) 75. Finally, we compared the measured similarity of  each network pair to random expectation: For each network, we performed 10 permutations of node indices, resulting in 100 permutations for a network pair which we used as random reference distribution to assess the measured overlap similarity. We then computed $z$-score and the corresponding empirical $p$-value. A network pair with $p-$value < 0.05 is considered significant. 


The MDS plot derived from Jaccard and Overlap Similarity is as follows:

```{r mds jaccard compute and plot}

pacman::p_load(ggrepel, MASS)

# load the precomputed data
if(!file.exists("../cache/network_jaccard_overlap_similarity_df.RDS")){
  source("../source/compute_jaccard_similarity.R")
} else{
  print("load pre-computed network similarity data")
  network_sim_df <- readRDS("../cache/network_jaccard_overlap_similarity_df.RDS")
}

# turn df to weight symmatrix matrix through graph
g_overlap <- graph_from_data_frame(network_sim_df[,c(1,2,4)] %>% 
                                     rename(., weight = overlapindex), directed = F)

sim_overlap <- get.adjacency(g_overlap, attr = "weight")

diag(sim_overlap) = 1

#change similarity to distance
dist_overlap = 1 - sim_overlap 

############
# MDS plot normal 

#### MDS plot for Kruskal
mds<- isoMDS(as.matrix(dist_overlap), k = 2)


# a data frame of MDS values
mds_df = data.frame(x = mds$points[,1], y = mds$points[,2], network = rownames(mds$points))

# add network metadata and node size
mds_df = mds_df %>%
  left_join(., g_prop_df %>% dplyr::filter(property=="Number of nodes") %>% 
              dplyr::select(network, value)) %>%
  left_join(., network_details) %>%
  dplyr::filter(!is.na(main_type)) %>%
  mutate(label = ifelse(!grepl("coex", network), subtype, "")
        # collabel = ifelse(!is.na(type), type, subtype)
         )

# plot the scatters of all networks
p <- mds_df %>% 
  ggplot() + 
  geom_point(aes(x, y, col = main_type, size = value), alpha = 0.5) + 
  geom_text_repel(aes(x, y, label = label)) + 
  theme_cowplot() +theme(
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.ticks.y=element_blank(),
    axis.text.y=element_blank()) + 
  xlab("MDS1") + ylab("MDS2") +
  scale_color_manual(values = c("#F8B100", "#005564"))+
  guides(col = F, size  = F)


p

#ggsave("../Figs/scatter_Network_complementarity_MDS_Overlap.pdf",plot = p, width = 4, height = 4)
```
### Summary statistics of the network similarity

Overlap index:

```{r}
network_sim_df %>% pull(overlapindex) %>% summary
```


```{r, include=FALSE}
# overlap between co-expression networks
median_coex_overlap <- network_sim_df %>% dplyr::filter(grepl("coex", V1), grepl("coex", V2)) %>% pull(overlapindex) %>% median

# overlap between non co-expression networks
median_noncoex_overlap <- network_sim_df %>% dplyr::filter(!grepl("coex", V1), !grepl("coex", V2)) %>% pull(overlapindex) %>% median
```
Median overlap indices between co-expression networks are `r median_coex_overlap` and non co-expression networks are `r median_noncoex_overlap`.


```{r, eval=FALSE, include=FALSE}

# this analysis aimed to quickly reconstruct the network similarity on graph format

library(ggraph)
library(tidygraph)

# full network connection
# network similarity
net_sim_core <- network_sim_df  %>% dplyr::filter(overlapindex > 0.01, !grepl("coex",V1)|grepl("core", V1), !grepl("coex",V2)|grepl("core", V2))

graph <- as_tbl_graph(net_sim_core) %>% 
    mutate(Popularity = centrality_degree())

# plot using ggraph
ggraph(graph, layout = 'kk') + 
    geom_edge_fan(aes(alpha = stat(index),  edge_width = overlapindex), show.legend = FALSE) + 
    geom_node_point(aes(size = Popularity)) + 
  geom_node_text(aes(label = name)) +
  guides(legend=FALSE)+
 #   facet_edges(~year) + 
    theme_graph(foreground = 'steelblue', fg_text_colour = 'white')
```

```{r, eval=FALSE, include=FALSE}
# represent as heatmap
net_sim_core <- network_sim_df  %>% dplyr::filter(!grepl("coex",V1)|grepl("core", V1), !grepl("coex",V2)|grepl("core", V2))

ggplot(net_sim_core) + geom_tile(aes(x = V1, y = V2, fill = (overlapindex))) + scale_fill_distiller(direction = -1)
```



```{r, heatmap_similarity_allnetwork_jaccard, include=TRUE, eval=TRUE}
## Jaccard heatmap

## remove coex_core from the map
p1 = network_sim_df[,1:3] %>% dplyr::filter(!grepl("core", V1), !grepl("core", V2)) %>% 
  # heatmap plot
  ggplot() + geom_tile(aes(x = V1, y = V2, fill = jaccardIndex)) + scale_fill_distiller(direction = 1) + xlab("") +ylab("") + ggtitle("Jaccard similarity among all networks") + theme_minimal()+ theme(axis.text.x = element_text(angle = 90))

p1
```
##  Global similarity (Suppl. Fig. 1g)

We use core transcriptional modules to represent all of the co-expression network. The heatmap below shows the Jaccard and Overlap similarity.

```{r heatmap_similarity_network_jaccard}
## remove coex_core from the map
considered_networks = c("coex_core", "reactome_copathway", "ppi", "MP", "HP", "GOMF", "GOBP")

labels = c("co-expression", "co-pathway", "PPI", "MP", "HP", "GOMF", "GOBP ")


# Jaccard index
#######
p1 = network_sim_df[,1:3] %>% dplyr::filter(V1 %in% considered_networks, V2 %in% considered_networks ) %>% 
  # rescale factor
  mutate(V1 = factor(V1, levels = considered_networks, labels = labels),
         V2 = factor(V2, levels = considered_networks, labels = labels)) %>%
  # heatmap plot
  ggplot() + geom_tile(aes(x = V1, y = V2, fill = jaccardIndex)) + scale_fill_distiller(direction = 1) + xlab("") +ylab("") + ggtitle("Jaccard similarity") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 1))

#ggsave("../Figs/heatmap_overlap_index_aggregated_jaccard.pdf", p1, width = 5, height = 4)

# Overlap index
#######
overlap_df <- network_sim_df[,c(1,2,4)] %>% dplyr::filter(V1 %in% considered_networks, V2 %in% considered_networks ) %>% 
  # rescale factor
  mutate(V1 = factor(V1, levels = considered_networks, labels = labels),
         V2 = factor(V2, levels = considered_networks, labels = labels))

  # heatmap plot
p2 = ggplot(overlap_df) + geom_tile(aes(x = V1, y = V2, fill = overlapindex)) + 
  scale_fill_distiller(direction = 1) + 
  xlab("") +ylab("") + 
  ggtitle("overlap similarity ") + 
  theme_cowplot() + 
  theme(axis.text.x = element_text(angle = 45, hjust=1), legend.position = "bottom")

#ggsave("../Figs/heatmap_overlap_index_aggregated_overlap.pdf", p2, width = 4, height = 5)

p1 + p2
```



```{r}
sprintf("average overlap index for all networks are %f", mean(overlap_df$overlapindex))
```

## Network similairty randomization (Suupl. Fig. 1h)

For a given pair of network $g_A \in G(V_A, E_A)$ and $g_B \in G(V_A, E_A)$, we computed edge similarity through overlap index:

$$S_{AB}=\dfrac{|E_A \cap E_B|}{\text{min}(|E_A|,|E_B|)}$$

Foe each network, we performed 10 permutations of node indices, resulting in 100 permutations for a network pair, wehere we obtained the reference distribution for their similarity. We then computed $z$-score and the corresponding empirical $p$-value. A network pair with $p-$value < 0.05 is considered significant (`../source/network_overlap_randomisation.R`) 


```{r load results}

pacman::p_load(tidyverse, patchwork, ggrepel, cowplot)

# Perform the randomisation, or load from cahe (recommend)
source("../source/network_overlap_randomisation.R")

# network similarity values
network_sim_df <- readRDS("../cache/network_jaccard_overlap_similarity_df.RDS") %>% 
  dplyr::filter(!grepl("core", V1), !grepl("core" , V2))

# edge counts
ecounts <- readRDS("../cache/network_complementarity_topological.RDS")%>% 
  dplyr::filter(property=="Number of edges") %>% 
  pull(value, name = network) 

# compute minimum edge size for each pair
network_sim_df$min_ecount <- apply(network_sim_df, 1, function(x) min(ecounts[x[1]], ecounts[x[2]]))

```


```{r process the results}

# process jaccard and overlap index
jaccard_index <- lapply(randomisation_overlap_result, function(x) x$intersect/x$union)
overlap_index <- lapply(1:length(randomisation_overlap_result), function(x) 
  randomisation_overlap_result[[x]][['intersect']]/network_sim_df$min_ecount[x])

# compute mean and sd, zscore and pvalue
network_sim_df <- network_sim_df %>% 
  mutate(jaccard.mean = sapply(jaccard_index, mean),
         jaccard.sd = sapply(jaccard_index, sd),
   
         overlap.mean = sapply(overlap_index, mean),
         overlap.sd = sapply(overlap_index, sd),
         
         jaccard.zscore = (jaccardIndex-jaccard.mean)/jaccard.sd,
          jaccard.pval = pnorm(jaccard.zscore, lower.tail = F),
               
          overlap.zscore = (overlapindex-overlap.mean)/overlap.sd,
          overlap.pval = pnorm(overlap.zscore, lower.tail = F)
  )


# label p values into classes
network_sim_df <- network_sim_df %>% 
  # make p values in groups
  mutate(adjusted_overlapPval = p.adjust(overlap.pval, method = "BH"),
         overlap.pval_level = cut(adjusted_overlapPval, 
                                  breaks = rev(c(1,5e-2, 1e-3, 1e-4, 1e-5, 0)), 
                                  labels = rev(c("ns","*","**","***","****")), 
                                  include.lowest = T, ordered_result = T),
         # compute whether the pair are from both co-expression, or non co-expressions
         type1 = grepl("coex", V1),
         type2 = grepl("coex", V2),
         pair_name = paste(str_remove(V1,"coex_|reactome_"), 
                           str_remove(V2,"coex_|reactome_"), sep = " - "),
         # label only if overlap score higher than 0.2
         pair_label = ifelse(overlapindex > 0.2, pair_name, ""),
         type_pair = factor(type1+type2, levels = 0:2, labels = c("non.coex - non.coex",
                                                                  "coex - non.coex",
                                                                  "coex - coex"))) %>%
  mutate(overlap.pval_level = factor(overlap.pval_level, levels = rev(levels(overlap.pval_level)))) 


network_sim_df %>% count(overlap.pval_level) %>% paged_table()
```

Despite their wide range of similarity scores, we found that 955 out of 990 network pairs (96.5%) are significantly more similar than random expectation.



```{r plotting}
# stable plot results
p_scatter = ggplot(network_sim_df, aes(x = overlapindex, 
                                       y = log2(overlapindex/overlap.mean))) +
  geom_point(aes(col = overlap.pval_level)) + 
  scale_colour_viridis_d(direction = -1) + 
  theme_minimal() +
  xlab(expression(Similarity~(S[AB]))) + labs(title = "Network pair similarity", col = "significance") + 
  ylab(expression(log[2](S[AB]/mu[S[AB]]))) 

# plot by type 
p_scatter_by_type <- p_scatter + 
  facet_grid(. ~ type_pair) + 
  geom_text_repel(aes(label = pair_label)) + 
  theme_cowplot() +
  theme(legend.position = "bottom")


# p value plot by level
pval_lv_count_df <- network_sim_df %>%  count(overlap.pval_level, name = "count")

pval_lv_count_by_type_df  <- network_sim_df %>%  
  count(overlap.pval_level, type_pair, name = "count") %>%
  group_by(type_pair) %>%
  mutate(prop = count/sum(count)) 
  

p_count_by_type = ggplot(pval_lv_count_by_type_df, aes(x = overlap.pval_level, y = prop)) + 
  geom_col(aes(fill = overlap.pval_level)) + 
  scale_fill_viridis_d(direction = -1) + 
  xlab("Significance") + ylab("proportion") + guides(fill = F) +
  theme_minimal() + facet_grid(. ~ type_pair) + labs(title = "Network similarity significance level") +
  theme_cowplot()


p = p_scatter_by_type/p_count_by_type

#ggsave("../Figs/scatter_randomisation_network_similarity.pdf", p, width = 7, height = 6)

p
```

Interestingly, we also observed that networks on different scales (i.e. among non co-expression layers) are all significantly similar, showing that there are key edges being maintained across genotype to phenotype.


