# Using differential modularity to contextualize rare disease gene clusters

A case study of microscopic analysis using our multiple networks framework. We explore the differential modularity of gastroenterological diseases as an example.


```{r computing specific info}

# load cached networks and diseases and LCC results
source("../source/cache_all_networks_and_disease_genes.R")

pacman::p_load(treemap, igraph, tidygraph, ggraph)

disease_of_interest <- "Rare genetic gastroenterological disease" 


# unique disease set under the term
individual_gene_set <- individual_diseases_genes %>% dplyr::filter(grepl(disease_of_interest, roots))

# all genes associated
genes_set <- rare_genetic_diseases_genes$disgene_list[[disease_of_interest]]

# list of all significant networks 
signif_nets <- processed_result_df %>% dplyr::filter(name == disease_of_interest, LCC.signif != "none") %>% arrange(desc(LCC.zscore)) %>% pull(network)

# degrees each gene in each network
degree_diseasegenes <- sapply(g[signif_nets], function(x) degree(x, v = intersect(V(x)$name, genes_set)))


# igraph object of significant networks for the disease group
disease_graph <- lapply(signif_nets, function(x) g[[x]]- setdiff(V(g[[x]])$name, genes_set))
names(disease_graph) <- signif_nets

```

## Individual disease gene association 

There are `r `nrow(individual_gene_set)` individual disease terms under `r disease_of_interest`, involving `r length(genes_set)` genes.

```{r}
# print the paged table
individual_gene_set %>% 
  select(label, genes, n_genes) %>% 
  arrange(-n_genes) %>%
  rmarkdown::paged_table()
```

## Treemap association (Fig. 4)

The tree map for descendants of `r disease_of_interest` are as follows (limited to terms with at least two associated genes).

```{r treemap}

# set a threshold for minimal amount of genes required to include in the plotting
threshold = 2

individual_gene_set_wrap <- individual_gene_set %>% dplyr::filter(n_genes >= threshold) %>% select(label, n_genes)

n_removed <- sum(individual_gene_set$n_genes<=threshold)
#individual_gene_set_wrap <- rbind(individual_gene_set_wrap, data.frame(label = sprintf("Other (%i diseases with one or two associated genes)", n_removed), n_genes = n_removed))


# Create data
{group <- ifelse(individual_gene_set_wrap$n_genes > threshold+2, 
                individual_gene_set_wrap$label, str_trunc(individual_gene_set_wrap$label, 50, side = "right"))

value <- individual_gene_set_wrap$n_genes
data <- data.frame(group,value)
 

treemap(data,
        palette = "Set3",   
        title = disease_of_interest,
            index="group",
            vSize="value",
            type="index"
            )
}

# treemap - uncomment to plot
#pdf("../Figs/treemap_gastroenterological.pdf", width = 5, height = 4)
```


The module significance computation shows the following networks being significant:

```{r}
processed_result_df %>% 
  dplyr::filter(name == disease_of_interest, LCC.signif != "none") %>% 
  arrange(desc(LCC.zscore)) %>%
  select(subtype, main_type, N_in_graph, LCC.size, LCC.mean, LCC.sd, LCC.zscore, correctedPval) %>%
  knitr::kable()
```

## Differential modularity (Fig. 4)

The network differential modularity can be visualised as follows:

```{r}
    # convert the igraph object into df for ggraph visualisation
disease_graph_df <- lapply(disease_graph, as_data_frame) %>% 
  bind_rows(., .id = "network") %>% 
  mutate(network = factor(network, levels = signif_nets))
    
    
disease_tidygraph <- as_tbl_graph(disease_graph_df) %>% 
  mutate(degree = centrality_degree(mode = 'all'))
    
tidygraph_ggrpah <- ggraph(disease_tidygraph, layout = 'kk')

 p <- tidygraph_ggrpah + 
          geom_edge_density(aes(fill = network),
                            show.legend = FALSE) + 
          geom_edge_fan(aes(alpha = stat(index), 
                            col = network), 
                          show.legend = FALSE) + 
          geom_node_point(col = "white",
                          alpha = 0.25,
                           show.legend  = F, 
                           size = 1) + 
          facet_edges(~factor(network, levels = signif_nets)) + 
          # black theme
            theme_graph(fg_text_colour = 'white', 
                      background = "black",
                      text_colour = "white") 
 
# uncomment when need to replot and save
#ggsave2(sprintf("../Figs/networks_%s.pdf", disease_of_interest), plot = p, width = 10, height = 6)
#ggsave2(sprintf("../Figs/networks_%s_separate_dimmed.png", disease_of_interest), plot = p, width = 10, height = 6) 

p
```

