# Modularity as quantification of pathobiological relevance

Note: this was the old version and the indexed one should be replaced by the new one.

```{r load data}
pacman::p_load(patchwork, cowplot, ggstatsplot)
knitr::opts_chunk$set(echo=FALSE)

if(!file.exists("../cache/fold_cv_processed_results_revisioned_processed.RDS")){
  source("../source/process_10fold_retrieval_for_revision.R")
}
AUC_folds_summary_updated_plot <- readRDS("../cache/fold_cv_processed_results_revisioned_processed.RDS")
```

## ROC-AUC comparison among all diseases and and network selection (Fig. 5b)
```{r}
palettes <- c('#66c2a5','#8da0cb',"#fec44f" , '#fc8d62')

p <- ggplot(AUC_folds_summary_updated_plot,aes(x=name, fill = label, group = label)) +
   geom_ribbon(aes(ymin = Q25, ymax = Q75), alpha = 0.25) +
  geom_line(aes(y = med, col = label), linetype = 2) +
  coord_flip() + xlab("disease") + ylab("fold AUC") + theme_minimal_hgrid() + theme(legend.position = "bottom") +
  ggtitle("ROC-AUC comparison", subtitle = "area spans 25, 50 and 75% quantile") +
  #scale_fill_brewer(type = "qual") + scale_color_brewer(type = "qual")
  scale_fill_manual(values=palettes) + scale_color_manual(values=palettes)

#ggsave("../Figs/AUC_folds_comparison_short_names_updated.pdf", p, width = 6, height = 6)

p
```


## Corresponding p-value based on statistical test for avg ROC curve (Fig. 5c)

```{r}
AUC_folds_summary_updated_plot2 <- AUC_folds_summary_updated_plot %>%
  mutate(label_plot = fct_recode(label, `Single most\n relevant layer`="SingleMostSignif"),
         label_plot = fct_relevel(label_plot, "All layers", after = 1))

palettes2 <- c('#8da0cb','#66c2a5',"#fec44f" , '#fc8d62')

p_compare_main <- ggwithinstats(
  data = AUC_folds_summary_updated_plot2,# %>% filter(!excluded_layer %in% c("CoEx", "CoEssential")) %>% ungroup,
  x = label_plot,
  y = med,
  p.adjust.method = "fdr",
  #title = "AUC values",
#  caption = "Data source: `WRS2` R package",
  ggtheme = theme_cowplot(), 
  xlab = "Network layers",
sample.size.label = F,
pairwise.display = "significant",
  ylab = "AUC", 
#ggplot.component = list(scale_color_brewer(palette = "Blues"), scale_fill_brewer(palette = "Blues")),
type = "non-parametric"#, 
#ggplot.component = c(col = "grey")
#  ggtheme = ggthemes::theme_fivethirtyeight()
) + scale_color_manual(values = palettes2)

#ggsave(plot = p_compare_main, filename = "../Figs/10fold_cv_method_compared_boxplot_updated.pdf", width = 6.5, height = 4, scale = 0.9)

p_compare_main
```



## relationship between the disease, number of signifiant networks, and number of genes (Fig. 5d)


```{r}
## load the LCC results 
source("../source/cache_all_networks_and_disease_genes.R")
## count number of networks, only coex

N_signif_net = processed_result_df %>% 
  filter(type=="Co-expression", LCC.signif!="none") %>%
  count(name, name = "significant_tissues")


## load the gene sets
Orphanet_df <-rare_genetic_diseases_genes$disgene_df %>% select(name, N)

Orphanet_df <- left_join(Orphanet_df, N_signif_net)

# retrieve median values
AUC_folds_perf_compare <- AUC_folds_summary_updated_plot %>% 
  pivot_wider(., id_cols = name, names_from = label, values_from = med) %>%
  mutate(dif = `Significant layers` - `All layers`) %>%
 # rename(name_short = "name") %>%
  left_join(x=., y = Orphanet_df) #, by = c(name = "name_short"))

AUC_allnets_median <- AUC_folds_summary_updated_plot %>% 
#    rename(name_short = "name") %>%
  dplyr::select(-Q25, -Q75) %>%
  dplyr::filter(label == 'Significant layers') %>%
  left_join(x=., y = Orphanet_df)


p1 <- ggplot(AUC_folds_perf_compare %>%
               ungroup %>%
               mutate(label = ifelse(dif< 0, name,""))
             , aes(x = significant_tissues, y = dif)) + 
  geom_point(col = "grey40") + 
  theme_cowplot() + guides(size = F) + 
  geom_hline(yintercept = 0, linetype = "dotted", col = "grey20") + 
 # stat_summary(fun.data= mean_cl_normal) + 
  geom_smooth(method='lm', se = F) +
  xlab("# Significant tissues") + ylab(expression("Performance diff. ("*Delta*"AUC)"))

p2 <- ggplot(AUC_allnets_median, aes(x = N, y = med)) + geom_point(col = "grey40") + theme_cowplot() + 
  guides(size = F) + xlab("# Associated genes") + ylab("Median AUC") + 
  scale_x_log10() +
   geom_smooth(method='lm', se = F)   



p <- p1 + p2

#ggsave(filename = "../Figs/scatter_AUC_vs_ntissues_updated.pdf",plot = p,  height = 3, width = 6)

p
```
### Corresponding statistical parameters on the left plot (Performance difference vs #tissue):

```{r}
cor.test(AUC_folds_perf_compare$dif,AUC_folds_perf_compare$significant_tissues, method = "spearman")
```

### Corresponding statistical parameters on the right plot (Performance vs #genes):
```{r}
cor.test(AUC_allnets_median$N, AUC_allnets_median$med, method = "spearman")
```

## Performance difference: using relevant networks vs all networks (Suppl. Fig. 7b)


```{r}
p <- AUC_folds_perf_compare %>%
  left_join(., distinct(AUC_folds_summary, name)# %>% rename(name_short = "name")
            , by="name") %>%
  ungroup %>% 
  mutate(name = fct_reorder(name, significant_tissues)) %>%
  ggplot(., aes(x = name, y = significant_tissues, fill=dif)) + geom_col() + theme_minimal() + coord_flip() + 
  scale_fill_gradient2(low = "#66c2a5", high = "#8da0cb") + 
  ylab("# Significant tissues") + xlab("") + labs(fill = "")+ ggtitle("Performance difference", subtitle = "Selected networks vs all networks")

#ggsave("../Figs/barchart_n_tissues_vs_performance_difference.pdf", height = 4, width = 6)

p
```




