# Application to gene prioritization in rare disease patients

## Number of candidate genes per patient
```{r}
gene_list <- read_csv("../data/patient_gene_list.csv") %>%
  mutate(Patient = factor(Patient, level = c("N_064", "N_126", "N_020", "N_026", "N_003", "N_129", "N_007", 
"N_040"), labels = paste0("P", 1:8)))


gene_list %>% count(Patient, name = "number of genes")
```

Average number of genes with high confident variants detected in each patient:

```{r}
nrow(gene_list)/length(unique(gene_list$Patient))
```


These patient are diagnosed, i.e. has a confirmed causal variant. We wonder whether the causal gene can be picked up by the algorithm. The patient phenotype belong to rare genetic neurological disorder. 

## Patient phenotypes (Suppl. Fig. 4c)
```{r}
patient_hpo_df <- read_csv("../data/patient_hpo_terms.csv", 
                           col_names = c('Patient', "hpo_id", "hpo_label"), skip = 1) %>%
  mutate(hpo_id = str_trim(hpo_id, side = 'both')) %>%
  mutate(patient_ID = factor(Patient, levels = unique(Patient), label = paste0("P", 1:length(unique(Patient)))))

hpo_counts <- patient_hpo_df %>%
  count(hpo_label) %>%
  arrange(n)  %>%
  mutate(hpo_label = factor(hpo_label, levels = hpo_label))

patient_hpo_df$hpo_label = factor(patient_hpo_df$hpo_label, levels = hpo_counts$hpo_label)

p = ggplot(patient_hpo_df %>% mutate(found=T), aes(x=patient_ID, y =hpo_label, fill = found)) + 
  #geom_tile(col = "white") + 
  geom_point(aes(colour = found), size = 6) +
  guides(fill = F, col=F) + 
  theme_minimal() + xlab("Patient") + ylab("HPO label") +
  coord_equal()

#ggsave("../Figs/heatmap_patient_phenotype.pdf", p)

p
```

## Patient-specific rankings

The prioritization was performed using informed propagation, with seed genes specific to each patient.

```{r}
precomputed_patient_ranked_genes <- "../cache/patient_ranking_results.RDS"

if(!file.exists(precomputed_patient_ranked_genes)){
  source("../source/patient_causal_gene_prediction.R")
}

# # load precomputed results
patient_annotated_df <- readRDS(precomputed_patient_ranked_genes) %>%
  mutate(Patient = factor(Patient, level = c("N_064", "N_126", "N_020", "N_026", "N_003", "N_129", "N_007", 
"N_040"), labels = paste0("P", 1:8)))
```


Below is the causal gene for each patient, and how well the algorithm performs for each case.

```{r}
# show the causal genes and how well the algorithm performs
correct_rank <- patient_annotated_df %>% 
  filter(Diagnostic=="YES", network_set == "signif") %>%
  select(-network_set, -avg) %>% 
  arrange(network_rank, -total_variants)

correct_rank
```


## Patient ranking plot (Fig. 5f, Suppl. Fig. 4d)

```{r patient_gene_ranking}
# ensure it keeps the order 

p_rank = list()
for( i in unique(patient_annotated_df$network_set)){
  correct_rank <- patient_annotated_df %>% 
  filter(Diagnostic=="YES", network_set == i) %>%
  select(-network_set, -avg) %>% 
  arrange(network_rank, -total_variants)
  
p_rank[[i]]  <- ggplot(correct_rank) +
  geom_segment( aes(x=Patient, xend=Patient, y=0, yend=total_variants), color="grey", size = 4) +
  geom_point(aes(x=Patient,y=network_rank), size=5, col="orange", shape = 15) +
  geom_text(aes(x=Patient,y=network_rank, label = network_rank), col = "white", fontface = "bold")+
  theme_light() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.border = element_blank()
  ) +
  xlab("Patient") +
  ylab("# Candidate genes")

#ggsave(p_rank[[i]], filename = paste0("../Figs/patient_gene_ranking_",i,".pdf"), width = 2.5, height = 3)
}

library(patchwork)
p_rank[["signif"]] + ggtitle("Signif") +
  p_rank[["all"]] + ggtitle("All networks") +  
  p_rank[["ppi"]]  + ggtitle("PPI only")
```

## Patient AUC plot (Fig. 5g)

```{r}
AUC_plot <- patient_annotated_df %>%
  mutate(pos = (network_rank-1)/(total_variants-1), 
         label = ifelse(Diagnostic=="NO", 0,1))

pacman::p_load(pROC)
rocvals <- list()
rocvals_df <- list()
for(i in unique(AUC_plot$network_set)){
  df <- AUC_plot %>% filter(network_set==i)
  response <- df$label
  predictor <- df$pos
  rocvals[[i]] <- pROC::roc(response, predictor)
  rocvals_df[[i]] <- tibble(Sensitivity =  rocvals[[i]]$sensitivities, Specificity = rocvals[[i]]$specificities)
}


rocvals_merged_df <- bind_rows(rocvals_df, .id = "rank_type")

palettes <- c('#66c2a5','#fc8d62','#8da0cb')

p <- ggplot(rocvals_merged_df, aes(x = Specificity, y = Sensitivity, col = rank_type)) + 
  geom_line(size=2) + 
  scale_x_reverse() + 
  scale_color_manual(values = palettes) + guides(col = FALSE) +
  cowplot::theme_cowplot()

#ggsave(p, filename = "../Figs/patient_gene_ranking_AUC.pdf", width = 3, height = 3)

p
```

The AUC for each method is as shown below.

```{r}
tibble(methods = names(rocvals), auc  = sapply(rocvals, function(x) x$auc))
```

The corresponding p-values are:

```{r}
pairs = combn(names(rocvals),2) %>% t %>% as.tibble()
auc_compare_test <- apply(pairs,1, function(x) roc.test(rocvals[[x[1]]], rocvals[[x[2]]], ))

pairs$pval <- sapply(auc_compare_test, function(x) x$p.value)

pairs
```


```{r old variant plot, eval=F, include=F}

# old variant plots, removed
dir_stored <- "../Figs/Patient_gene_prioritisation/"
dir.create(dir_stored)
for(i in unique(variants_annotated$Patient)){
   df <- variants_annotated %>% 
     dplyr::filter(Patient == i) %>% 
     mutate(true_call =  ifelse(Diagnostic == "NO", F, T))
   p = ggplot(df, aes(x = LocalRank, y = avg_prob, col = true_call, label = GENE)) + geom_point() + ggrepel::geom_label_repel() + ggtitle(i)
   ggsave(paste0(dir_stored, i, ".pdf"), p)
}



dir.create(dir_stored)
for(i in unique(variants_annotated$Patient)){
   df <- variants_annotated %>% 
     dplyr::filter(Patient == i) %>% 
     mutate(true_call =  ifelse(Diagnostic == "NO", F, T)) %>%
     select(GENE, LocalRank, true_call) %>%
     right_join(., rank_df_disease, by = c("GENE"="GeneName")) %>%
     mutate(genelab = ifelse(!is.na(LocalRank), GENE, ""),
            col_lab = ifelse(!is.na(true_call), true_call, "not in patient"))
   p = ggplot(df, aes(x = GlobalRank, y = avg_prob, col = true_call, label = genelab)) + geom_point() + ggrepel::geom_label_repel() + ggtitle(i)
   ggsave(paste0(dir_stored, "global_", i, ".pdf"), p)
}

write_csv(x = variants_annotated, file = paste0(dir_stored, "avg_prob_patients.csv"))
```



