---
title: "Supplementary Codes"
description: |
 Reproducible walk-through of codes and data for Buphamalai et.al.
author:
  - first_name: "Pisanu"
    last_name: "Buphamalai"
    affiliation: Max Perutz Labs | CeMM
    orcid_id: 0000-0002-9024-4193
    affiliation_url: https://www.maxperutzlabs.ac.at/  
  - first_name: "Tomislav"
    last_name: "Kokotovic"
    affiliation: Neurology, MedUni Vienna | LBI-RUD | CeMM
  - first_name: "Vanja"
    last_name: "Nagy"
    affiliation: Neurology, MedUni Vienna | LBI-RUD | CeMM
  - first_name: "Jörg"
    last_name: "Menche"
    affiliation: Max Perutz Labs | Mathematics, Uni Vienna | CeMM
    orcid_id: 0000-0002-1583-6404
date: "`r Sys.Date()`"
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
    code_folding: true 
---

# Introduction

This supplementary report is aimed to be a reproducible walk-through guide for figures and analyses complementing the manuscript: Buphamalai et.al., *Network analysis reveals rare disease signatures across multiple levels of biological organization*, submitted to Nature Communications. PLease find the following guidelines:

1. The appearance of sections in this document is at the same order as in the manuscript, and can be navigated using the Table of Content (ToC) appeared on the top left corner of this document, with subsections corresponding to exact figures/statistics 
2. The corresponding code chunks to each figures/analyses are provided above the output, and are hidden by default. To expand each code chunk, click `Show code`.
3. This report mainly contains visualization and post-processing of major analyses. Heavier computations were pre-computed, and corresponding `R`, `sh`, or `py` scripts required for each analysis are mentioned for each section. These files are available in `./source` folder.
4. The pre-computed results are saved in `./cache` folder and can be downloaded from: [link](https://drive.google.com/file/d/1T7tJMojIbELeT-aLOD_Pv639eUgdGqUh/view) 
5. The corresponding `Rmd` files used to produce this report can be found in `./report/` folder. The report mainly focuses on reproducing results shown in the manuscripts. For explanation, discussion, and references, please refer to the main manuscript.

Note: genome-phenome data used in the analysis in `7.1_Prioritize_ID_RDconnect.Rmd` is subject to controlled access by RD-connect. 

```{r global setup, include=FALSE}
filter <- dplyr::filter
select <- dplyr::select
```


```{r child = '1_Network_complementarity.Rmd'}
```

```{r child = '1.1_coex_nodes_edges_characterization.Rmd '}
```

```{r child = '1.3_Assortativity_and_PPI_subsets.Rmd'}
```

```{r child = '1.2_literature_bias_analyses.Rmd'}
```

```{r child = '2_Co-expression_network_characterisation.Rmd'}
```

```{r child = '3_OrphaNet_Rare_Disease_processing.Rmd'}
```

```{r child = '4_Quantifying_network_modularity.Rmd'}
```

```{r child = '5_contextualization.Rmd'}
```

```{r child = '6_Cross_validation_retrieval.Rmd'}
```

```{r child = '6.1_Performance_layer_removal.Rmd'}
```

```{r child = '7_Patient_gene_prioritization.Rmd'}
```

```{r child = '7.1_Prioritize_ID_RDconnect.Rmd'}
```


# Session Information

```{r}
pacman::p_load(report)
report_session <- report(sessionInfo())
write_lines(report_session, "report_session.md")
print(report_session)
```

